---
title: "Revisiting the von Bertalanffy Seasonal Cessational Growth Function of Pauly et al. (1992)"
author: "Derek H. Ogle"
date: "April 29, 2016"
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 4.5
  html_document:
    fig_height: 3.5
    fig_width: 4.5
csl: american-fisheries-society.csl
bibliography: SeasonalGrowth.bib
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(34738422)  # for reproducibility
library(FSA)
library(FSAsim)
library(latex2exp)
library(knitr)

source("iCalc_tpr.R")
source("VBSCGF.R")

# setup figure, table, and equation captioning
library(captioner)
figcaps <- captioner(prefix="Figure")
tblcaps <- captioner(prefix="Table")
eqncaps <- captioner(prefix="Equation")
eqncaps("vbMain")
eqncaps("vbMain2")
eqncaps("SomersMod")
eqncaps("PaulyMod")
eqncaps("PaulyMod2")

vbSO <- vbFuns("Somers2")
```

# Introduction

The mean length-at-age for many animals (e.g., CITATIONS) is often modeled with the von Bertalanffy growth function (VBGF; @vonBertalanffy1938).  The parameterization of the VBGF attributable to @BevertonHolt1957 is most common and may be expressed as 

$$ L_{t} = L_{\infty}(1-e^{-q}) \quad \quad \text{`r paste0("(",eqncaps("vbMain",display="num"),")")`} $$

with 

$$ q=K(t-t_{0}) \quad \quad \text{`r paste0("(",eqncaps("vbMain2",display="num"),")")`} $$

where $L(t)$ is the expected or average length at time (or age) $t$, $L_{\infty}$ is the asymptotic mean length, $K$ is a measure of the exponential rate of approach to the asymptote [@SchnuteFournier1980], and $t_{0}$ is the theoretical time or age (generally negative) at which the mean length would be zero.

Many animals exhibit seasonal oscillations in growth as a response to seasonal changes in environmental factors such as temperature, light, and food supply (CITATIONS).  `r eqncaps("vbMain2",display="cite")` of the traditional VBGF has been modified, usually with a sine function, to model these seasonal oscillations in growth.  The most popular of these modifications, from @HoenigChoudarayHanumara1982 and @Somers1988, is

$$
\begin{aligned}
q= & K(t-t_{0}) \\
   & +\frac{CK}{2\pi}sin(2\pi(t-t_{s}))-\frac{CK}{2\pi}sin(2\pi(t_{0}-t_{s}))  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("SomersMod",display="num"),")")`}
\end{aligned}  
$$

where $C$ modulates the amplitude of the growth oscillations and corresponds to the proportional decrease in growth at the depth of the oscillation (i.e., "winter"), and $t_{s}$ is the time between time 0 and the start of the convex portion of the first sinusoidal growth oscillation (i.e., the inflection point).  If $C$=0, then there is no seasonal oscillation and `r eqncaps("SomersMod",display="cite")` reduces to  `r eqncaps("vbMain2",display="cite")` and the typical VBGF (`r figcaps("SomersExPlot",display="cite")`).  If $C$=1, then growth completely stops once a year at the "winter-point" ($WP$), whereas values of 0&lt;$C$&lt;1 result in reduced, but not stopped, growth during the winter (`r figcaps("SomersExPlot",display="cite")`).  Some confusion has surrounded this function, though @GarciaBerthouetal2012 carefully clarified its form.

\ 

```{r echo=FALSE, results="hide"}
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.4,0),tcl=-0.2,las=1)
WP <- 0.55; Linf <- 30; K <- 0.3; t0 <- -0.1
C <- c(0,0.5,1,2)
clr <- col2rgbt(rep("black",4),1/(1:4))
t <- seq(0,3,length.out=299)
plot(vbSO(t,Linf=c(Linf,K,t0,C=C[1],WP))~t,type="l",lwd=2,col=clr[1],
     ylim=c(0,20),ylab="Mean Length",xlab="Age (years)",xaxt="n")
axis(1,0:3)
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[2],WP))~t,lwd=2,col=clr[2])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[3],WP))~t,lwd=2,col=clr[3])
lines(vbSO(t,Linf=c(Linf,K,t0,C=C[4],WP))~t,lwd=2,col=clr[4])

legend("topleft",paste("C",C,sep="="),lwd=2,col=clr,bty="n",inset=-0.02)

figcaps("SomersExPlot",paste0("Example VBGF using ",eqncaps('SomersMod',display='cite')," with $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", $t_{s}$=",WP-0.5," (or $WP$=",WP,"), and four different values of $C$."))
```

`r figcaps("SomersExPlot")`

\ 

Values of $C$&gt;1 (or &lt;0) in  `r eqncaps("SomersMod",display="cite")` allow seasonal decreases in mean length-at-age (`r figcaps("SomersExPlot",display="cite")`).  A decrease in mean length is unlikely for organisms whose skeletons largely preclude shrinkage [@Paulyetal1992], although a seasonal decrease in mean length-at-age is possible if size-dependent overwinter mortality occurs [@GarciaBerthouetal2012].  @Paulyetal1992 modified `r eqncaps("SomersMod",display="cite")` to include a true seasonal no-growth period where mean length was not allowed to decrease and that included a smooth transition of the modeled mean length-at-age into and out of the no-growth period.  Specifically, their modification is

\ 

$$
\begin{aligned}
q= & K^{'}(t^{'}-t_{0}) \\
   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t^{'}-t_{s})\right)-\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}(t_{0}-t_{s})\right) \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod",display="num"),")")`}
\end{aligned}  
$$

\ 

\noindent where $NGT$ is the "no-growth time" or the length of the no growth period (as a fraction of a year) and $t^{'}$ is found by "subtracting from the real age ($t$) the total no-growth time occurring up to age $t$" [@Paulyetal1992].  Furthermore,  @Paulyetal1992 noted that the units of $K$ changed from $year^{-1}$ in `r eqncaps("SomersMod",display="cite")` to $(1-NGT)^{-1}$ in `r eqncaps("PaulyMod",display="cite")`.  To eliminate confusion, they suggested using $K^{'}$ in `r eqncaps("PaulyMod",display="cite")`, as we do here.

@Paulyetal1992 devised  `r eqncaps("PaulyMod",display="cite")` by assuming $C$=1 (i.e., that the rate of growth is 0 at the $WP$) and replacing $2\pi$ with $\frac{2\pi}{1-NGT}$ (i.e., restricting the oscillation to the growth period and noting that $K^{'}$ only operates during the growth period).  Their modification may be described geometrically (though not algorithmically) in two steps.  First, the seasonal growth function from with $C$=1 in `r eqncaps("SomersMod",display="cite")` is fit to the observed lengths and ages that have had the cumulative $NGT$ subtracted (i.e., using $t^{'}$).  The growth trajectory is then separated at each $WP$ and horizontal segments that are $NGT$ units long are inserted at these points.  This forms a growth trajectory that smoothly transitions into and out of the no-growth periods (`r figcaps("PaulyExPlot",display="cite")`).

\ 

```{r echo=FALSE, results="hide"}
## Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
WP <- 0.55; NGT <- 0.3
Linf <- 30; Kpr <- 0.4; t0 <- 0
Pcf <- c(Linf,Kpr,t0,WP,NGT)
## Get growth trajectory
t <- seq(-0.1,3.3,length.out=499)
PL <- VBSCGF(t,Linf=Pcf)
## Calculate WPs and start and end of each NGT
# find WPs and Lengths at WPs
# find start and end of NGT
## Make schematic plot
plot(c(-1,-1),xlim=c(-0.1,3.5),ylim=c(0,22),xaxt="n",
     ylab="Length",xlab="Age (years)")
## Put on NGT polygons
WPs <- 0:2+WP
LatWPs <- VBSCGF(WPs,Linf=Pcf)
SNGTs <- WPs-NGT/2
ENGTs <- WPs+NGT/2
for (i in 1:length(SNGTs)) polygon(c(SNGTs[i],SNGTs[i],ENGTs[i],ENGTs[i]),c(0,LatWPs[i],LatWPs[i],0),
                                   col=col2rgbt("black",1/20),border=NA)
## Mark the WPs
arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
## Mark the NGTs
arrows(SNGTs,LatWPs+1.2,ENGTs,LatWPs+1.2,lwd=2,length=0.025,angle=90,code=3)
## Put on the growth trajectory
lines(PL~t,lwd=2)
## Add some important values to the age axis
axis(1,0:3)
axis(1,SNGTs,tcl=-0.2)
axis(1,ENGTs,tcl=-0.2)
## Add the t-prime labels
# makes inside ticks
axis(1,at=c(0:3,SNGTs,ENGTs),labels=NA,tcl=0.2)
axis(1,at=1:3,labels=iCalc_tpr(1:3,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=SNGTs,labels=iCalc_tpr(SNGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=ENGTs,labels=iCalc_tpr(ENGTs,WP,NGT),tcl=0.2,line=-1.7,lwd=0)
axis(1,at=3.3,labels="<= t",tick=FALSE)
axis(1,at=3.3,labels="<= t'",tick=FALSE,line=-1.7,lwd=0)

figcaps("PaulyExPlot",paste0("Example VBGF using ",eqncaps('PaulyMod',display='cite')," or ",eqncaps('PaulyMod2',display='cite')," with $L_{\\infty}$=",Linf,", $K^{'}$=",formatC(Kpr,format="f",digits=4),", $t_{0}$=",t0,", $NGT$=",NGT,", and $t_{s}$=",WP-0.5," or $WP$=",WP,".  Each $WP$ is shown by the arrow and each no-growth period is shown by the horizontal error bar centered on the $WP$ arrow and by the gray region that extends to the x-axis.  The ages adjusted for the $NGT$ (i.e., $t^{'}$) are shown above the x-axis."))
```

`r figcaps("PaulyExPlot")`

\ 

@Paulyetal1992 provided a "diskette" that contained a computer program to estimate the parameters of `r eqncaps("PaulyMod",display="cite")`.  The diskette is difficult (at best) to obtain and the source code is no longer available (D. Pauly, pers. comm.).  @Paulyetal1992 did describe the operations performed by their program, but there is no description of how $t^{'}$ was operationalized.  This is an important step in using `r eqncaps("PaulyMod",display="cite")` because $t^{'}$ is a function of $t$, but it is also a function of $NGT$ and $t_{s}$, which are parameters to be estimated during the model-fitting process.  Thus, the values for $t^{'}$ change with each iteration of the non-linear model-fitting process.

Therefore, the objectives of this note are to (i) describe a slight modification of `r eqncaps("PaulyMod",display="cite")` that eases the calculation of $t^{'}$ while also providing a more meaningful parameter; (ii) operationalize the calculation of $t^{'}$ for use in model fitting; (iii) provide an (open-source) algorithm for the modified function for use in model fitting; and (iv) demonstrate the fitting of the modified function to length-at-age data.

# The Modified Seasonal Cessation Growth Function

In `r eqncaps("SomersMod",display="cite")`, $WP=t_{s}+\frac{1}{2}$ because the sine function has a period (i.e., the growth period) of one year.  Because the growth period is compressed in `r eqncaps("PaulyMod",display="cite")` to be $1-NGT$, the start of the no-growth period in `r eqncaps("PaulyMod",display="cite")` is $t_{s}+\frac{1-NGT}{2}$.  The center of the no-growth period is the start of the no-growth period plus half of the no-growth time, or

$$ t_{s}+\frac{1-NGT}{2}+\frac{NGT}{2} = t_{s}+\frac{1}{2} = WP $$.

Thus, $WP=t_{s}+\frac{1}{2}$ is the center of the no-growth period in `r eqncaps("PaulyMod",display="cite")`.

By simple substitution, `r eqncaps("PaulyMod",display="cite")` may be modified to include $WP$ rather than $t_{s}$.

\ 

$$
\begin{aligned}
q= & K^{'}(t^{'}-t_{0}) \\
   & +\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t^{'}-WP-\frac{1}{2}\right)\right)  \quad \quad \quad \quad \text{`r paste0("(",eqncaps("PaulyMod2",display="num"),")")`} \\
   & -\frac{K^{'}(1-NGT)}{2\pi}sin\left(\frac{2\pi}{1-NGT}\left(t_{0}-WP-\frac{1}{2}\right)\right)
\end{aligned}  
$$

\ 

As noted by @Paulyetal1992 the calculation of $t^{'}$ depends on the observed age ($t$) and the cumulative no-growth time prior to $t$.  In practice, the calculation of $t^{'}$ also depends on the position of the no-growth period within a year.  Here, the position of the no-growth period is defined by $SNG$ and $NGT$.  However, because the $SNG$ has a one-to-one relation with $WP$ and $t_{s}$, the no-growth period can also be defined relative to these parameters.

The following algorithm is used to convert from observed ages ($t$) to ages adjusted for cumulative $NGT$ prior to age $t$ ($t^{'}$).

1. Shift the age ($t$) by subtracting the start of the no-growth period ($SNG$) from $t$, such that a whole number age represents the start of a no-growth period.  For example, if $SNG$=0.4, then $t$=2.4 will become 2.0 and $t$=2.9 will become 2.5.
1. Subtract the whole number age from the shifted age such that the remaining decimal represents the fraction of a shifted year.  For example, a 0 will result if the shifted age is 2.0 and a 0.5 will result if the shifted age is 2.5.
1. Substract the $NGT$ from the value from the previous step.
1. If the value from the previous step is negative, then the age is within the no-growth period and the negative value should be replaced with a zero.  Otherwise, the positive value represents the fraction of the growth period completed.
1. Add the value from the previous step to the total growth time completed (i.e., the product of the number of growth periods completed and the length of the growth period ($1-NGT$)).
1. Compute $t{'}$ by adding back the $SNG$ that was subtracted in Step 1.

Further examples of $t^{'}$ values relative to $t$ values are shown in `r figcaps("PaulyExPlot",display="cite")`.  This algorithm for computing $t^{'}$ is implemented in an R [@R2016] function as shown in Appendix 1.  With this, `r eqncaps("PaulyMod2",display="cite")` is easly implemented as an R function as shown in Appendix 2.

# Fitting the Modified Model
Simple simulated data were initially used to demonstrate the fitting of `r eqncaps("PaulyMod2",display="cite")`.  The data were simulated by randomly selecting 200 real numbers from a uniform distribution between 0 and 5 to serve as observed ages, plugging these ages into `r eqncaps("PaulyMod2",display="cite")` to compute a mean length for each age, and then adding a random deviate from a normal distribution with a mean of 0 and a standard deviation of $\sigma$ to each mean length to simulate observed individual lengths.  The `nls()` function from R was then used to estimate parameter values for `r eqncaps("PaulyMod2",display="cite")` from the simulated data.  The "port" algorithm in `nls()` was used as $L_{\infty}$ and $K^{'}$ were constrained to be positive and $WP$ and $NGT$ were constrained to be between 0 and 1.  One would expect the parameter estimates to be very close to the values used to create the simulated data, as `r eqncaps("PaulyMod2",display="cite")` was used to both create and fit the data.

One example of a model fit to data that were simulated to have a "long" NGT, "late" WP, and low individual variability is shown in `r figcaps("PaulySimRes",display="cite")`.  In this example, the parameter estimates are quite close to the parameter values used to simulate the data (the set parameter values are within the confidence intervals for each parameter).


```{r echo=FALSE, results='hide'}
## Set some common parameters
Linf <- 30; K <- 0.3; t0 <- -0.1
## Set bounds for model-fitting
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
```
```{r echo=FALSE, results="hide", message=FALSE}
## Set some specific growth parameters
# fairly typical north temperate situation
WP <- 0.8; NGT <- 0.5
sigma <- 0.3
ages <- runif(200,min=0,max=5)
lens <- VBSCGF(ages,Linf,K,t0,WP,NGT)+rnorm(200,0,sigma)
dB1 <- data.frame(ages,lens)
## Try fitting the VBSCGF model
sv <- list(Linf=32,K=0.32,t0=0,WP=0.9,NGT=0.3)
fitB1 <- nls(lens~VBSCGF(ages,Linf,K,t0,WP,NGT),data=dB1,
             start=sv,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
cfs <- cbind(Est=coef(fitB1),confint(fitB1))
## A plot
# Setup
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
plot(lens~ages,data=dB1,pch=19,col=col2rgbt("black",1/3),
     xlab="Age (years)",ylab="Length",xlim=c(0,5),ylim=c(0,18))
curve(VBSCGF(x,cfs[,"Est"]),from=0,to=6,lwd=2,add=TRUE)
WPs <- 0:5+cfs["WP","Est"]
LatWPs <- VBSCGF(WPs,Linf=cfs[,"Est"])
arrows(WPs,LatWPs+2,WPs,LatWPs+0.2,lwd=2,length=0.1)
SNGTs <- WPs-cfs["NGT","Est"]/2
ENGTs <- WPs+cfs["NGT","Est"]/2
arrows(SNGTs,LatWPs+1.2,ENGTs,LatWPs+1.2,lwd=2,length=0.025,angle=90,code=3)

legend("topleft",c(TeX("\\mathbf{Set Values}"),
                   TeX(paste("$L_{\\infty}$",Linf,sep=" = ")),
                   paste("K'",K,sep=" = "),
                   TeX(paste("$t_{0}",t0,sep=" = ")),
                   paste("WP",WP,sep=" = "),
                   paste("NGT",NGT,sep=" = ")),
        bty="n",inset=-0.02,cex=0.8)
legend("bottomright",c(TeX("\\mathbf{Estimated Values}"),
                       TeX(paste0("$L_{\\infty}$"," = ",
                                  formatC(cfs["Linf","Est"],format="f",digits=1),
                                  " (",formatC(cfs["Linf","2.5%"],format="f",digits=1),
                                  ", ",formatC(cfs["Linf","97.5%"],format="f",digits=1),")")),
                   paste0("K' = ",formatC(cfs["K","Est"],format="f",digits=2),
                                  " (",formatC(cfs["K","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["K","97.5%"],format="f",digits=2),")"),
                   TeX(paste0("$t_{0}$"," = ",
                                  formatC(cfs["t0","Est"],format="f",digits=2),
                                  " (",formatC(cfs["t0","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["t0","97.5%"],format="f",digits=2),")")),
                   paste0("WP = ",formatC(cfs["WP","Est"],format="f",digits=2),
                                  " (",formatC(cfs["WP","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["WP","97.5%"],format="f",digits=2),")"),
                   paste0("NGT = ",formatC(cfs["NGT","Est"],format="f",digits=2),
                                  " (",formatC(cfs["NGT","2.5%"],format="f",digits=2),
                                  ", ",formatC(cfs["NGT","97.5%"],format="f",digits=2),")")),
        bty="n",cex=0.8)

figcaps("PaulySimRes",paste0("Example VBGF using ",eqncaps('PaulyMod2',display='cite')," with parameters shown in the upper-left corner and $\\sigma$=",sigma,".  Parameter estimates (and 95% confidence intervals) from non-linear regression are shown the lower-right corner.  Each WP is shown by the arrow and each NGT is shown by the horizontal interval bar centered on the WP arrow."))
```

`r figcaps("PaulySimRes")`

\ 

`r eqncaps("PaulyMod2",display="cite")` was fit to simulated data sets that consisted of all combinations of "early" ($WP=0.2$) and "late" ($WP=0.8$) winter points, "short" ($NGT=0.2$) and "long" ($NGT=0.5$) no-growth periods, and "low" ($\sigma=0.3$) and "high" ($\sigma=0.5$) individual variabilities.  These limited simulations suggest that all parameters are consistently well estimated, with the possible exception of $t_{0}$.

\ 

```{r echo=FALSE, results="hide", message=FALSE}
WPs <- rep(c(0.8,0.2),each=4)
NGTs <- rep(rep(c(0.5,0.2),each=2),2)
sigmas <- rep(c(0.3,0.5),4)
ages <- runif(200,min=0,max=5)
sv <- list(Linf=32,K=0.32,t0=0,WP=0.5,NGT=0.5)

## Start with results from above
res <- matrix(nrow=8,ncol=18)
res[1,] <- c(WPs[1],NGTs[1],sigmas[1],round(cfs["WP",],2),round(cfs["NGT",],2),round(cfs["Linf",],1),round(cfs["K",],2),round(cfs["t0",],2))
for (i in 2:length(WPs)) {
  tmp <- data.frame(ages,lens=VBSCGF(ages,Linf,K,t0,WPs[i],NGTs[i])+rnorm(200,0,sigmas[i]))
  tmpfit <- nls(lens~VBSCGF(ages,Linf,K,t0,WP,NGT),data=tmp,
                start=sv,lower=lwrbnd,upper=uprbnd,
                algorithm="port",control=list(maxiter=100))
  tmpcfs <- cbind(Est=coef(tmpfit),confint(tmpfit))
  res[i,] <- c(WPs[i],NGTs[i],sigmas[i],round(tmpcfs["WP",],2),round(tmpcfs["NGT",],2),round(tmpcfs["Linf",],1),round(tmpcfs["K",],2),round(tmpcfs["t0",],2))
}
colnames(res) <- c("WPset","NGTset","sigmaset","WPest","WPlci","WPuci","NGTest","NGTlci","NGTuci","Linfest","Linflci","Linfuci","Kest","Klci","Kuci","t0est","t0lci","t0uci")

tblcaps("PaulySimRes",paste0("Parameter estimates from ",eqncaps('SomersMod',display='cite')," for simulated data using $L_{\\infty}$=",Linf,", $K$=",K,", $t_{0}$=",t0,", and varying values of $WP$, $NGT$, and $\\sigma$."))
```

`r tblcaps("PaulySimRes")`

```{r echo=FALSE, comment=""}
res
```

\ 



## Real Data

@Stewartetal2013 examined the growth of 215 Bonito (*Sarda australis*) sampled from commercial landings.  Detailed methods are described in @Stewartetal2013, but note here that fork lengths (mm) were measured for each fish and ages were the decimal age calculated as the number of opaque zones observed on otolith thin sections plus the proportion of the year after the designated birthdate.  @Stewartetal2013 fit `r eqncaps("SomersMod",display="cite")` to these data but constrained $C$ to not exceed 1.  Their model fit resulted in the boundary condition of $C=1$, which suggested that Bonito ceased to grow at *at least* one point.  As $C$ is at least equal to 1, this result also suggests that `r eqncaps("PaulyMod2",display="cite")` should be fit to these data to determine the length of the no-growth period [@Paulyetal1992].  Thus, `r eqncaps("PaulyMod2",display="cite")` is fit to these data here as an example.

```{r echo=FALSE, results='hide', message=FALSE}
## Bonito data
bon <- read.csv("Bonito.csv")

################################################################################
## Purely Demonstration
## See if results match that from Stewart et al. (2013)
vbSO <- vbFuns("Somers")
lwrbnd <- c(Linf=0,K=0,t0=-Inf,C=0,ts=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,C=1,ts=1)
svBon <- list(Linf=72,K=0.27,t0=0,C=0.6,ts=0.2)
fitBon1 <- nls(FL~vbSO(Age,Linf,K,t0,C,ts),data=bon,
               start=svBon,lower=lwrbnd,upper=uprbnd,
               algorithm="port",control=list(maxiter=100))
## Linf=71.9, K=0.27, t0=-1.92, C=1, and ts=0.09 ... they all match (within rounding)
cfBon1 <- cbind(Est=coef(fitBon1),confint(fitBon1))
################################################################################
```

```{r echo=FALSE, results='hide', message=FALSE}
## Fit new model to Bonito data
lwrbnd <- c(Linf=0,K=0,t0=-Inf,WP=0,NGT=0)
uprbnd <- c(Linf=Inf,K=Inf,t0=Inf,WP=1,NGT=1)
svBon <- list(Linf=72,K=0.27,t0=0,WP=0.5,NGT=0.2)
fitBon2 <- nls(FL~VBSCGF(Age,Linf,K,t0,WP,NGT),data=bon,
             start=svBon,lower=lwrbnd,upper=uprbnd,
             algorithm="port",control=list(maxiter=100))
cfBon2 <- cbind(Est=coef(fitBon2),confint(fitBon2))
```

`r eqncaps("PaulyMod2",display="cite")` fit the Bonito data slightly better than `r eqncaps("SomersMod",display="cite")` with slightly lower residual sums-of-squares (RSS) and Akaike Information Criterion (AIC) values.  The length of the no-growth period was estimated to be `r formatC(cfBon2["NGT","Est"],format="f",digits=3)` or `r formatC(cfBon2["NGT","Est"]*100,format="f",digits=1)`% of the year.  The $WP$ parameter from `r eqncaps("PaulyMod2",display="cite")` was equal to the $t_{s}$ parameter from `r eqncaps("SomersMod",display="cite")` plus 0.5.  The $L_\infty$ parameters were very similar between the two models, but the $t_{0}$ parameters differed somewhat.  Finally, $K$ from `r eqncaps("SomersMod",display="cite")` was equal to $K'$ from `r eqncaps("PaulyMod2",display="cite")` divided by $1-NGT$.

```{r echo=FALSE, results="hide"}
## A plot
par(xaxs="i",yaxs="i",mar=c(3,3,0.6,0.6),mgp=c(1.7,.5,0),tcl=-0.3,las=1)
plot(FL~Age,data=bon,pch=19,col="black",
     xlab="Age (years)",ylab="Fork Length (mm)",xlim=c(0,4),ylim=c(0,70))
curve(VBSCGF(x,coef(fitBon2)),from=0,to=4,lwd=4,add=TRUE)
curve(vbSO(x,coef(fitBon1)),from=0,to=4,lwd=2,add=TRUE,col="gray50")

figcaps("BonitoRes",paste0("Fork length at age for Australian Bonito with the best-fit of ",eqncaps('PaulyMod2',display='cite')," (black line) and ",eqncaps('SomersMod',display='cite')," (gray line) superimposed.  The parameter estimates (and 95% confidence intervals) from non-linear regression for ",eqncaps('PaulyMod2',display='cite')," are shown the lower-right corner."))
```

`r figcaps("BonitoRes")`

\ 

```{r echo=FALSE, results="hide"}
cfBon1 <- rbind(cfBon1,c(sum(summary(fitBon1)$residuals^2),NA,NA),c(AIC(fitBon1),NA,NA))
cfBon2 <- rbind(cfBon2,c(sum(summary(fitBon2)$residuals^2),NA,NA),c(AIC(fitBon2),NA,NA))
rownames(cfBon1)[6:7] <- rownames(cfBon2)[6:7] <- c("RSS","AIC")

tblcaps("BonitoRes",paste0("Parameter estimates, residual sums-of-squares (RSS), and Akaike Information Criterion (AIC) from the fits of ",eqncaps('SomersMod',display='cite')," and ",eqncaps('PaulyMod2',display='cite')," to the Bonito data."))
```

`r tblcaps("BonitoRes")`

```{r echo=FALSE, comment=""}
cat("Somers Model (Equation 3) fit")
print(round(cfBon1,2),na.print="-")
cat("Modified Pauly Model (Equation 6) fit")
print(round(cfBon2,2),na.print="-")
```

\ 

# Discussion

* General -- fits perfectly simulated data well.
* Assumptions -- WP same time each year and age, NGT same length each year and age
* Model-Fitting -- bounding parameters, more difficult to fit five parameters
* Practical Difference -- not large unless C>>1 and NGT>>0

# Appendices
## Appendix 1

```{r echo=FALSE, comment=""}
cat(readLines("iCalc_tpr.R"),sep="\n")
```

## Appendix 2

```{r echo=FALSE, comment=""}
cat(readLines("VBSCGF.R"),sep="\n")
```

# References
